---
title: "Assignment 2 ST464/ST684"
author: "AKSHAY MISHRA"
date: "`r format(Sys.time(), '%X %d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=4, fig.height=4)
```


```{r, eval=T, echo=FALSE}
suppressMessages(library(tidyverse))
suppressMessages(library(ggplot2))
suppressMessages(library(GGally))

```



#### Question 1



```{r eval=T} 
# change to eval=T
bfat <- read.table("bodyfat.txt", header=T)
bfat <- bfat[,c("neck","abdomen", "knee", "ankle")]
source("h2code.R")
```

(a)
```{r}
pairs(bfat, gap=0.5)
```

Yes, there is a presence of outliers in the figure at the neck, abdomen, knee and ankle plots.

(b)
```{r}
sapply(bfat[,-5],var)
p <- prcomp(bfat[,-5],scale=T,rext=T)
summary(p)
screeplot(p)
```

The percentage of variability for PC1 is 66.29 while the percetage of variability for PC2 in data set is 85.35

Most of the variability is covered by the first two components PC1 and PC2.
The graph becomes less steep after the second component.Therefore we take PC1 and PC2 into consideration.

(c)
```{r}
biplot(p,sacle=0,cex=c(.5,.5),cex.axis=.5)
```

Abdomen, neck and knee are highly correlated due to their closeness to each other. Rest for ankle case its away from the other three so not much of high correlation
Body fat values on upper right quadrant are high in PC1 and PC2, therefore high abdomen, neck and knee values and low ankle values.
  Body fat values on lower right quadrant are high in PC1 and low on PC2, therefore mostly have low values on abdomen, neck and knee and high ankle values.
  Body fat values on upper left quadrant have high values in PC2 and low on PC1, therefore have high values on abdomen, neck and knee and low value on ankle.
  Body fat values on lower left quadrant have low values in PC1 and PC2, therefore mostly have low values on abdomen, neck and knee and high ankle values.
  
There is a presence of outliers from the graph plotted.
The values are 84 and 31.

(d)
```{r}
p1 <- prcomp(bfat[-31,-84],sacle=T)
summary(p1)
biplot(p1,scale=0,cex=c(.5,.5),cex.axis=.5)
```

No visible outliers are there and variance for PC2 has declined.